---
title: "Logistic Regression predicting Student Dropout"
author: "Adnan Mayof"
date: "2024-02-02"
output: html_document
---

This R Markdown document guides through a logistic regression analysis for predicting student dropouts using R. Using student demographics, academic performance, and socio-economic data, the goal is to develop a robust predictive model. This model can assist educational institutions in resource allocation, implementing early interventions, and supporting at-risk students. The code covers data loading, exploration, and model building, while also introducing techniques like confusion matrices and ROC curves to enhance predictive accuracy assessment.

```{r}
knitr::opts_chunk$set(echo = TRUE)
```
 
```{r}
# Load Libraries
#install.packages("caTools")
#install.packages("ROCR")
library(caTools)
library(ROCR)
```



```{r}
# Step 1: Read Data
# Read the CSV file into R_Data
R_Data <- read.csv("R_Data.csv")
```

```{r}
# Step 2: Attach Dataframe
# Attach the R_Data dataframe for easier variable reference
attach(R_Data)
```


```{r}
# Step 3: Check Column Names
# Check Column Names
names(R_Data)
```

```{r}
# Step 4: Verify Data Structure
# Verify Data Structure
str(R_Data)
```

```{r}
# Step 5: Split Data
# Split the dataset into training and testing sets
split <- sample.split(R_Data, SplitRatio = 0.8)
split 
```

```{r}
# Step 6: Create Training and Testing Datasets
# Create training and testing datasets based on the split
training <- subset(R_Data, split == TRUE)
testing <- subset(R_Data, split == FALSE)
```

```{r}
# Step 7: Display First 5 Rows
# Display the first 5 rows of the training dataset
# A: training dataset
head(training, n = 5)

# Display the first 5 rows of the testing dataset
# B: testing dataset
head(testing, n = 5)
```


```{r}
# Step 8: Build Logistic Regression Models - Model 1
# Model1A with training dataset
model1A <- glm(Dropout ~ ., training, family = "binomial")
summary(model1A)

# Model1B with testing dataset
model1B <- glm(Dropout ~ ., testing, family = "binomial")
summary(model1B)
```

```{r}
# Step 8: Build Logistic Regression Models - Model 2
# ModelA2 with training dataset
modelA2 <- glm(Dropout ~ Marital_Status + Application_Mode - Application_Order, training, family = "binomial")
summary(modelA2)

# ModelB2 with testing dataset
modelB2 <- glm(Dropout ~ Marital_Status + Application_Mode - Application_Order, testing, family = "binomial")
summary(modelB2)
```

```{r}
# Step 8: Build Logistic Regression Models - Model 2
# ModelA2 with training dataset
modelA2 <- glm(Dropout ~ Marital_Status + Application_Mode - Application_Order, training, family = "binomial")
summary(modelA2)

# ModelB2 with testing dataset
modelB2 <- glm(Dropout ~ Marital_Status + Application_Mode - Application_Order, testing, family = "binomial")
summary(modelB2)
```

```{r}
# Step 8: Build Logistic Regression Models - Model 3
# ModelA3 with training dataset
modelA3 <- glm(Dropout ~ Application_Mode + Tuition_fees_up_to_date + Age_At_Enrollment + Course + Scholarship_Holder + Curricular_Units_2nd_Sem_Grade, training, family = "binomial")
summary(modelA3)

# ModelB3 with testing dataset
modelB3 <- glm(Dropout ~ Application_Mode + Tuition_fees_up_to_date + Age_At_Enrollment + Course + Scholarship_Holder + Curricular_Units_2nd_Sem_Grade, testing, family = "binomial")
summary(modelB3)
```


```{r}
# Step 9: Make Predictions
# A: training dataset
res_train <- predict(modelA3, training, type = "response")
head(res_train, n = 10)

# B: testing dataset
res_test <- predict(modelB3, testing, type = "response")
head(res_test, n = 10)
```

```{r}
# Step 10: Create Confusion Matrix
# A: training dataset
conf_matrix_train <- table(actualValue = training$Dropout, PredictedValue = res_train > 0.5)
conf_matrix_train

# B: testing dataset
conf_matrix_test <- table(actualValue = testing$Dropout, PredictedValue = res_test > 0.5)
conf_matrix_test
```

```{r}
# Step 11: Calculate Accuracy
# A: training dataset
accuracy_train <- sum(diag(conf_matrix_train)) / sum(conf_matrix_train)
accuracy_train

# B: testing dataset
accuracy_test <- sum(diag(conf_matrix_test)) / sum(conf_matrix_test)
accuracy_test
```

```{r}
# Step 12: Calculate Threshold
# A: training dataset
threshold_conf_matrix_train <- table(actualValue = training$Dropout, PredictedValue = res_train > 0.5)
threshold_accuracy_train <- sum(diag(threshold_conf_matrix_train)) / sum(threshold_conf_matrix_train)
threshold_accuracy_train

# B: testing dataset
threshold_conf_matrix_test <- table(actualValue = testing$Dropout, PredictedValue = res_test > 0.5)
threshold_accuracy_test <- sum(diag(threshold_conf_matrix_test)) / sum(threshold_conf_matrix_test)
threshold_accuracy_test
```


```{r}
# Step 13: Choose New Threshold using ROC Curve
# A: training dataset
ROCRPred_train <- prediction(res_train, training$Dropout)
ROCRPref_train <- performance(ROCRPred_train, "tpr", "fpr")
plot(ROCRPref_train, colorize = TRUE, print.cutoffs.at = seq(0.1, by = 0.1))

# B: testing dataset
ROCRPred_test <- prediction(res_test, testing$Dropout)
ROCRPref_test <- performance(ROCRPred_test, "tpr", "fpr")
plot(ROCRPref_test, colorize = TRUE, print.cutoffs.at = seq(0.1, by = 0.1))
```

```{r}
# Step 14: Choose New Threshold based on ROC Curve
# A: training dataset
new_threshold_conf_matrix_train <- table(actualValue = training$Dropout, PredictedValue = res_train > 0.3)
new_threshold_accuracy_train <- sum(diag(new_threshold_conf_matrix_train)) / sum(new_threshold_conf_matrix_train)
new_threshold_accuracy_train

# B: testing dataset
new_threshold_conf_matrix_test <- table(actualValue = testing$Dropout, PredictedValue = res_test > 0.3)
new_threshold_accuracy_test <- sum(diag(new_threshold_conf_matrix_test)) / sum(new_threshold_conf_matrix_test)
new_threshold_accuracy_test
```
 